---
title: "Neurological condition analysis"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "htmls") })
---

```{r setup, message=FALSE, warning=FALSE, results='hide'}
my_packages <- c('tidyverse', 'readxl', 'DT', 'rcartocolor', 'gghighlight', 
                 'cowplot', 'broom', 'devtools', 'tableone')

existing_pkgs <- installed.packages()[,"Package"]
to_install <- setdiff(my_packages, existing_pkgs)
if (length(to_install)) install.packages(to_install)
if (!'icd' %in% existing_pkgs) devtools::install_github('jackwasey/icd')

lapply(my_packages, library, character.only = TRUE)
theme_set(theme_bw() + 
            theme(legend.title = element_blank(),
                  panel.grid.minor = element_blank()))
for (r_file in list.files('R/', full.names = TRUE)) source(r_file)
set.seed(1)
```

```{r penn, message=FALSE}
mysite = 'penn'
mask_thres = 10 # count mask threshold
blur_abs = 0 # absolute max of blurring range

clin_raw <-
 read_csv(
   '../thrombotic-penn/data/LocalPatientClinicalCourse.csv',
   col_types = list(patient_num = col_character())
 )
demo_raw <-
 read_csv(
   '../thrombotic-penn/data/LocalPatientSummary.csv',
   col_types = list(patient_num = col_character()),
   na = '1900-01-01'
 )
obs_raw <-
 read_csv(
   '../thrombotic-penn/data/LocalPatientObservations.csv',
   col_types = list(patient_num = col_character())
 )
```


```{r nwu, message=FALSE}
# mysite = 'nwu'
# mask_thres = 0 # count mask threshold
# blur_abs = 0 # absolute max of blurring range

# clin_raw <- read_csv(
#   '../../Phase2.1/Date_Shifted_Files/LocalPatientClinicalCourse_Shift.csv',
#   col_types = list(patient_num = col_character()))
# demo_raw <- read_csv(
#   '../../Phase2.1/Date_Shifted_Files/LocalPatientSummary_Shift.csv',
#   col_types = list(patient_num = col_character()),
#   na = '1900-01-01')
# obs_raw <- read_csv('../../Phase2.1/LocalPatientObservations.csv',
#                     col_types = list(patient_num = col_character()))
```


```{r message=FALSE, warning=FALSE}
loinc <- read_csv('public-data/loinc-map.csv')
neuro_icds_10 <- read_excel('public-data/2020-09-10_neuro-icd10.xlsx') %>% 
  rename('icd' = `ICD-10`)
```

## Data pre-processing

In this initial analysis, we will consider neuro patients as those with a neuro code on the day **of** or **anytime after** their first day of admission.
These patients developed neurological conditions after COVID-19 related admission.

Only examining ICD-10 for now.

*Note: We will need to discuss whether we want to include codes assigned on the day of admission as comorbidities.*


```{r}
neuro_patients <- obs_raw %>% 
  filter(
    days_since_admission >= 0,
    concept_code %in% neuro_icds_10$icd) %>% 
  distinct(patient_num, concept_code)

neuro_pt_post <- unique(neuro_patients$patient_num)
  
non_neuro_patients <- data.frame(
  patient_num = setdiff(demo_raw$patient_num, neuro_pt_post)) %>% 
  mutate(concept_code = 'NN')

readmissions <- clin_raw %>% 
  group_by(patient_num) %>% 
  mutate(delta_hospitalized = diff(c(in_hospital[1], in_hospital))) %>% 
  ungroup() %>% 
  filter(delta_hospitalized != 0,
         in_hospital == 1) %>%
  add_count(patient_num, name = 'n_readmissions') %>%
  arrange(desc(n_readmissions))

length(unique(readmissions$patient_num))

re_admin_stats <- readmissions %>% 
  select(patient_num, n_readmissions) %>% 
  distinct() %>% 
  count(n_readmissions)

re_admin_stats

re_admin_stats %>% 
  ggplot() +
  geom_col(aes(x = n_readmissions, y = n))
```

Length of stay is estimated based on the last day of having an "observation".

```{r}
days_count_min_max <- obs_raw %>%
  group_by(patient_num) %>%
  summarise(
    distinct_days = n_distinct(days_since_admission),
    min_hos = min(days_since_admission),
    # n_stay = max(days_since_admission),
    .groups = 'drop'
  )

demo_df <- demo_raw %>%
  mutate(time_to_severe = severe_date - admission_date,
         time_to_severe = ifelse(time_to_severe < 0, NA, time_to_severe),
         time_to_death = death_date - admission_date,
         time_to_death = ifelse(time_to_death < 0, NA, time_to_death),
         readmitted = patient_num %in% readmissions$patient_num,
         neuro_post = patient_num %in% neuro_pt_post %>% 
           as.factor() %>% 
           fct_recode(`Has neurological condition` = "TRUE",
                      `No neurological condition` = "FALSE"),
         Survival = as.factor(deceased) %>% 
           fct_recode(Alive = "0", Deceased = "1"),
         sex = as.factor(sex),
         race = as.factor(race),
         age_group = as.factor(age_group),
         Severity = as.factor(severe) %>% 
           fct_recode(Severe = "1", `Non-severe` = "0"),
         n_stay = as.numeric(last_discharge_date - admission_date,
                             units="days")) %>% 
  left_join(days_count_min_max, by = 'patient_num')
```

## Demographics (Table 1)

```{r}
formatted_demo <- demo_df %>% 
  select(Sex = sex, `Age group` = age_group, 
         Race = race, Severity, Survival, 
         `Time to severity onset (days)` = time_to_severe, 
         `Time to death (days)` = time_to_death, 
         Readmitted = readmitted, neuro_post)

## export table 1 to csv
table_one_vars <- setdiff(colnames(formatted_demo), 'neuro_post')
demo_table <- tableone::CreateTableOne(
  data = formatted_demo,
  vars = table_one_vars,
  strata= "neuro_post")

table_one <- print(
  demo_table,
  showAllLevels = TRUE,
  formatOptions = list(big.mark = ","),
  printToggle = FALSE
)

## print table 1 html
table1::table1(
  ~ Sex + `Age group` + Race + Severity + Survival + Readmitted +
    `Time to severity onset (days)` + `Time to death (days)` 
  | neuro_post,
  data = formatted_demo,
  overall = "Total",
  export = "results/demographics.csv")
```

## Comorbidity

```{r}
n_patients <- nrow(demo_raw)
obs_raw <- obs_raw %>% 
  filter(concept_type %in% c("DIAG-ICD10", "DIAG-ICD9"))
table(demo_df$neuro_post)
```

New variable `neuro_post` to indicate whether the patient developed neurological conditions after COVID-19 related admission.

## Map comorbidities

List of comorbidity abbreviations for reference purposes,
then run the `map_char_elix_codes()` function:

```{r}
# for elixhauser
comorb_names_elix <- get_quan_elix_names()
comorbs_elix <- as.vector(comorb_names_elix$Abbreviation)

# t1: earliest time point to consider comorbidities
# t2: latest time point to consider comorbidities
# example <- t1 = -365, and t2 = -1 will map all all codes up to a year prior but before admission (admission = day 0)

comorb_elix <- map_char_elix_codes(
  df = obs_raw,
  comorb_names = comorb_names_elix,
  t1 = -365,
  t2 = -15,
  map_type = 'elixhauser'
)
```

`map_char_elix_codes()` outputs a 2-element lists: 

<!-- 1- Dataframe indicating which comorbidities were mapped -->

1- Dataframe of the Charlson/Elixhauser comorbidity index scores

2- Dataframe of the specific mapped codes arranged by frequency.
Comorbidity mapping of all patients with ICD codes prior to admission, regardless of whether or not they had a comorbidity.

```{r}
# elixhauser
index_scores_elix <- comorb_elix$index_scores %>% 
  rename('elixhauser_score' = van_walraven_score)
# van Walraven is a modification of Elixhauser comorbidity measure
# doi.org/10.1097/MLR.0b013e31819432e5
mapped_codes_table_elix <- comorb_elix$mapped_codes_table
comorb_names_elix$Abbreviation <- as.character(comorb_names_elix$Abbreviation)
```

Here we can determine how many unique patients we have comorbidity index for prior to admission:
```{r}
# elixhauser
length(unique(index_scores_elix$patient_num))
comorb_unique <- select(index_scores_elix, patient_num, elixhauser_score)
```

## Elixhauser comorbidity matrix

Heatmap of number of patients in each comorbidity type:

```{r}
n_comorbs <- colSums(index_scores_elix[, comorbs_elix])
pos_comorbs <- names(n_comorbs[n_comorbs > 0])
index_scores_elix[, pos_comorbs] %>% 
  cor() %>% 
  heatmap(cexRow = 0.5, cexCol = 0.5)
```

### Elixhauser Distribution

```{r}
comorb_unique %>% 
  left_join(demo_df, by = 'patient_num') %>% 
  mutate(severe = as.factor(severe) %>% 
           fct_recode('Severe' = '1', 'Non-severe' = '0')) %>% 
  ggplot(aes(x = neuro_post, y = elixhauser_score, fill = neuro_post)) + 
  geom_boxplot(outlier.colour = NA) + 
  geom_jitter(alpha = 0.2, width = 0.2, size = 0.6) +
  guides(fill = FALSE) +
  labs(x = NULL) +
  NULL
```

### Elixhauser Comorbidity Table

```{r}
concat <- function(x, y){
  paste0(x, ' (', round(y, 3)*100, '%)')
}

scores_unique <- left_join(index_scores_elix, demo_df, by = 'patient_num')

scores_neuro <- obs_raw %>% 
  # 1 patient can have different code but each only counted once
  distinct(patient_num, concept_code) %>% 
  left_join(neuro_icds_10, by = c('concept_code' = 'icd')) %>% 
  left_join(scores_unique, by = 'patient_num') %>% 
  filter(!is.na(elixhauser_score)) %>%
  mutate(concept_code = case_when(
           is.na(`Neurological Disease Category`) ~ 'NN',
           TRUE ~ concept_code) %>% 
           as.factor() %>% 
           fct_reorder(- elixhauser_score),
         `Neurological Disease Category` =
           as.factor(`Neurological Disease Category`) %>% 
           fct_reorder(elixhauser_score)) %>% 
  {.}

combined_table1_elix <- get_table1(
  index_scores_elix %>% filter(patient_num %in% neuro_pt_post),
  comorbidities = comorb_names_elix,
  comorbidities_map = comorb_names_elix$Abbreviation) %>% 
  rename('n_neuro_pats' = n_patients) %>% 
  left_join(get_table1(index_scores_elix,
                       comorbidities_map = comorb_names_elix$Abbreviation, 
                       comorbidities = comorb_names_elix), 
            by = c("Comorbidity", "Abbreviation"))

## apply blurring and masking
blur_vec <- matrix(sample(seq(-blur_abs, blur_abs),
                   4*length(comorbs_elix),
                   replace = TRUE), ncol = 4)

combined_table1_elix <- combined_table1_elix %>%
  mutate(
    n_patients = ifelse(n_patients < mask_thres, 0, n_patients + blur_vec[, 1]),
    n_neuro_pats = ifelse(n_neuro_pats < mask_thres, 0, n_neuro_pats + blur_vec[, 2]),
    n_non_neuro = n_patients - n_neuro_pats,
    n_non_neuro = ifelse(abs(n_non_neuro) < mask_thres, 0, n_non_neuro + blur_vec[, 3]),
    prop_patients = n_patients / nrow(demo_raw),
    prop_neuro_pats = n_neuro_pats / nrow(demo_raw),
    prop_non_neuro = n_non_neuro / nrow(demo_raw)
  ) %>%
  arrange(desc(n_neuro_pats))

combined_table1_elix %>% 
  transmute(
    Comorbidity, Abbreviation,
    `Has neurological condition` = concat(n_neuro_pats, prop_neuro_pats),
    `No neurological condition` = concat(n_non_neuro, prop_non_neuro),
    Total = concat(n_patients, prop_patients)) %>% 
  datatable() 

combined_table1_elix %>% 
  pivot_longer(c(n_patients, n_neuro_pats)) %>% 
  mutate(fullname = glue::glue('{Comorbidity} ({Abbreviation})'))%>% 
  ggplot(aes(x = value, y = fct_reorder(fullname, value), fill = name)) +
  geom_col(position = 'dodge') +
  scale_fill_brewer(palette = 'Dark2', direction = -1,
                    guide = guide_legend(reverse=TRUE),
                    label = c('Neuro', 'All')) +
  labs(y = NULL)
```

## Short term outcomes

Do hospitalized COVID-19 patients with neurological conditions have a different short-term natural history (hospital days, death, severity, readmission) from those without?

Resolution: Elixhauser score vs comorbidity type, Neurological condition (TRUE/FALSE) vs. neurological ICD code?

### Severity, Death, Length of stay

```{r}
severe_reg_elix <- glm(severe ~ neuro_post + elixhauser_score + sex + age_group + race,
   data = scores_unique, family = 'binomial') %>% 
  tidy()
severe_reg_elix
deceased_reg_elix <- glm(deceased ~ neuro_post + elixhauser_score + sex + age_group + race,
   data = scores_unique, family = 'binomial') %>% 
  tidy()
deceased_reg_elix

n_stay_reg_elix <- lm(n_stay ~ neuro_post + elixhauser_score + race + sex + age_group, 
   data = scores_unique) %>%
  tidy()

n_stay_reg_elix
```

**Elixhauser**

```{r}
nstay_df <- neuro_patients %>% 
  bind_rows(non_neuro_patients) %>% 
  left_join(demo_df, by = 'patient_num') %>%
  left_join(index_scores_elix, by = 'patient_num') %>%
  # mutate(concept_code = fct_reorder(concept_code, n_stay)) %>% 
  left_join(neuro_icds_10, by = c('concept_code' = 'icd')) %>%
  mutate(full_icd = case_when(
    concept_code == 'NN' ~ 'No neurological condition',
    TRUE ~ paste0(`ICD-10 Description`, ' (', concept_code, ')')) %>%
             as.factor() %>% fct_reorder(n_stay)) 

summarised_stay <- nstay_df %>%
  group_by(concept_code) %>% 
  summarise(mean_stay = mean(n_stay),
            median_stay = median(n_stay),
            sd_stay = sd(n_stay),
            mean_elix = mean(elixhauser_score, na.rm = TRUE),
            median_elix = median(elixhauser_score, na.rm = TRUE),
            sd_elix = sd(elixhauser_score, na.rm = TRUE),
            n_patients = n(),
            n_patients = ifelse(n_patients < mask_thres, 0, n_patients + blur_vec),
            prop_deceased = mean(deceased),
            prop_severe = mean(severe),
            .groups = 'drop')

```

```{r save-results}
list_results = list(
  summarised_stay = summarised_stay, 
  n_stay_reg_elix = n_stay_reg_elix, 
  severe_reg_elix = severe_reg_elix, 
  deceased_reg_elix = deceased_reg_elix, 
  # table_one = table_one,
  combined_table1_elix = combined_table1_elix)

site_results <- paste0(mysite, '_results')
assign(site_results, list_results)
save(list = site_results,
     file = paste0('results/', mysite, '-results.rda'))

# write.csv(table_one, file = 'results/penn-table1.csv')
# write.csv(combined_table1_elix,
#           paste0('results/', mysite, '-combined-elix.csv'), 
#           row.names = FALSE)
```

```{r nwu-save-results}
# summarised_stay_nwu = summarised_stay
# n_stay_reg_nwu = n_stay_reg
# severe_reg_nwu = severe_reg
# deceased_reg_nwu = deceased_reg
# table_one_nwu = table_one
# write.csv(table_one_nwu, file = "results/nwu-table1.csv")
# 
# save(summarised_stay_nwu,
#      severe_reg_nwu,
#      deceased_reg_nwu,
#      table_one_nwu,
#      n_stay_reg_nwu, 
#      file = 'results/nwu-results.rda')
```

