---
title: "Neurological condition analysis"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "htmls") })
---

```{r setup, message=FALSE, warning=FALSE, results='hide'}
my_packages <- c('tidyverse', 'readxl', 'DT', 'rcartocolor', 'gghighlight', 
                 'cowplot', 'broom', 'devtools', 'tableone')

existing_pkgs <- installed.packages()[,"Package"]
to_install <- setdiff(my_packages, existing_pkgs)
if (length(to_install)) install.packages(to_install)
if (!'icd' %in% existing_pkgs) devtools::install_github('jackwasey/icd')

lapply(my_packages, library, character.only = TRUE)
theme_set(theme_bw() + 
            theme(legend.title = element_blank(),
                  panel.grid.minor = element_blank()))
for (r_file in list.files('R', full.names = TRUE)) source(r_file)
```

## Data read-in

Change this following chunk of code for your site-specific data:

```{r penn, message=FALSE}
# set.seed(1)
mysite = 'penn'
mask_thres = 10 # count mask threshold
blur_abs = 0 # absolute max of blurring range
clin_raw <-
read_csv(
  '../thrombotic-penn/data/LocalPatientClinicalCourse.csv',
  col_types = list(patient_num = col_character())
)
demo_raw <-
read_csv(
  '../thrombotic-penn/data/LocalPatientSummary.csv',
  col_types = list(patient_num = col_character()),
  na = '1900-01-01'
)
obs_raw <-
read_csv(
  '../thrombotic-penn/data/LocalPatientObservations.csv',
  col_types = list(patient_num = col_character())
)
```


```{r nwu, message=FALSE}
# mysite = 'nwu'
# mask_thres = 0 # count mask threshold
# blur_abs = 0 # absolute max of blurring range
# 
# clin_raw <- read_csv(
#    '../../Phase2.1/Date_Shifted_Files/LocalPatientClinicalCourse_Shift.csv',
#    col_types = list(patient_num = col_character()))
# demo_raw <- read_csv(
#    '../../Phase2.1/Date_Shifted_Files/LocalPatientSummary_Shift.csv',
#    col_types = list(patient_num = col_character()),
#    na = '1900-01-01')
# obs_raw <- read_csv('../../Phase2.1/LocalPatientObservations.csv',
#                      col_types = list(patient_num = col_character()))
```


```{r message=FALSE, warning=FALSE}
# loinc <- read_csv('public-data/loinc-map.csv')
neuro_icds_10 <-
  read_excel('public-data/2020-09-10_neuro-icd10_CNSvPNS.xlsx') %>%
  rename('icd' = `ICD-10`,
         'pns_cns' = `Nervous system Involvement (1=central, 2=peripheral)`) %>%
  mutate(pns_cns = as.factor(pns_cns) %>% fct_recode(Central = '1',
                                                     Peripheral = '2'))

neuro_icds_10[, -(1:2)] %>% 
  write_csv('results/neuro_icds_10_pcns.csv')
```

## Data pre-processing

In this initial analysis, we will consider neuro patients as those with a neuro code on the day **of** or **anytime after** their first day of admission.
These patients developed neurological conditions after a COVID-19 related admission.

We're only examining ICD-10 for now.

*Note: comorbidities are estimated with ICD codes recorded from 365-15 days prior to the day of admission.*

```{r}
neuro_patients <- obs_raw %>% 
  filter(
    days_since_admission >= 0,
    concept_code %in% neuro_icds_10$icd) %>% 
  distinct(patient_num, concept_code)

neuro_pt_post <- unique(neuro_patients$patient_num)
  
non_neuro_patients <- data.frame(
  patient_num = setdiff(demo_raw$patient_num, neuro_pt_post)) %>% 
  mutate(concept_code = 'NN')

readmissions <- clin_raw %>% 
  group_by(patient_num) %>% 
  mutate(delta_hospitalized = diff(c(in_hospital[1], in_hospital))) %>% 
  ungroup() %>% 
  filter(delta_hospitalized != 0,
         in_hospital == 1) %>%
  add_count(patient_num, name = 'n_readmissions') %>%
  arrange(desc(n_readmissions)) %>% 
  select(patient_num, n_readmissions) %>% 
  distinct()

nrow(readmissions)

readmissions %>% 
  count(n_readmissions) %>% 
  blur_it('n', blur_abs, mask_thres) %>% 
  ggplot() +
  geom_col(aes(x = n_readmissions, y = n))
```


*Note*: Length of stay is calculated as the period length from admission date to the last discharge date.

```{r}
days_count_min_max <- obs_raw %>%
  group_by(patient_num) %>%
  summarise(
    distinct_days = n_distinct(days_since_admission),
    min_hos = min(days_since_admission),
    .groups = 'drop'
  )

demo_df <- demo_raw %>%
  mutate(time_to_severe = severe_date - admission_date,
         time_to_severe = ifelse(time_to_severe < 0, NA, time_to_severe),
         time_to_death = death_date - admission_date,
         time_to_death = ifelse(time_to_death < 0, NA, time_to_death),
         readmitted = patient_num %in% readmissions$patient_num,
         neuro_post = patient_num %in% neuro_pt_post %>% 
           as.factor() %>% 
           fct_recode(neuro_cond = "TRUE",
                      no_neuro_cond = "FALSE"),
         Survival = as.factor(deceased) %>% 
           fct_recode(Alive = "0", Deceased = "1"),
         sex = as.factor(sex),
         race = as.factor(race),
         age_group = as.factor(age_group),
         Severity = as.factor(severe) %>% 
           fct_recode(Severe = "1", `Non-severe` = "0"),
         n_stay = as.numeric(last_discharge_date - admission_date,
                             units="days")) %>% 
  left_join(days_count_min_max, by = 'patient_num')
```

## Demographics (Table 1)

### Overall
```{r}
vars_to_obfs <- c("sex", 'age_group', "race", "Severity",
                  "Survival", "readmitted")
get_stats <- function(x) demo_stats(demo_df, x, blur_abs, mask_thres)
demo_obfus_table <- lapply(vars_to_obfs, get_stats) %>% 
  do.call(rbind, .)
datatable(select(demo_obfus_table, variable, contains('pres')))
```

### Other descriptive stats

Length of stay, number of readmissions, time to severity onset, time to death

```{r}
nstay_obfus_table <- nstay_stats(demo_df, blur_abs, mask_thres)
severity_obfus_table <- severity_stats(demo_df, blur_abs, mask_thres)
survival_obfus_table <- survival_stats(demo_df, blur_abs, mask_thres)
readmission_obfus_table <- demo_df %>% 
  left_join(readmissions, by = 'patient_num') %>% 
  replace_na(list(n_readmissions = 0)) %>% 
  readmission_stats(blur_abs, mask_thres)

```

## Comorbidity

```{r}
n_patients <- nrow(demo_raw)
obs_raw <- obs_raw %>% 
  filter(concept_type %in% c("DIAG-ICD10", "DIAG-ICD9"))
table(demo_df$neuro_post)
```

New variable `neuro_post` to indicate whether the patient developed neurological conditions after COVID-19 related admission.

## Map comorbidities

List of comorbidity abbreviations for reference purposes,
then run the `map_char_elix_codes()` function:

```{r}
# for elixhauser
comorb_names_elix <- get_quan_elix_names()
comorbs_elix <- as.vector(comorb_names_elix$Abbreviation)

# t1: earliest time point to consider comorbidities
# t2: latest time point to consider comorbidities
# example <- t1 = -365, and t2 = -1 will map all all codes up to a year prior but before admission (admission = day 0)

comorb_elix <- map_char_elix_codes(
  df = obs_raw,
  comorb_names = comorb_names_elix,
  t1 = -365,
  t2 = -15,
  map_type = 'elixhauser'
)
```

`map_char_elix_codes()` outputs a 2-element lists: 

<!-- 1- Dataframe indicating which comorbidities were mapped -->

1- Dataframe of the Charlson/Elixhauser comorbidity index scores

2- Dataframe of the specific mapped codes arranged by frequency.
Comorbidity mapping of all patients with ICD codes prior to admission, regardless of whether or not they had a comorbidity.

```{r}
# elixhauser
index_scores_elix <- comorb_elix$index_scores %>% 
  rename('elixhauser_score' = van_walraven_score)
# van Walraven is a modification of Elixhauser comorbidity measure
# doi.org/10.1097/MLR.0b013e31819432e5
mapped_codes_table_elix <- comorb_elix$mapped_codes_table
comorb_names_elix$Abbreviation <- as.character(comorb_names_elix$Abbreviation)
```

Here we can determine how many unique patients we have comorbidity index for prior to admission:
```{r}
comorb_unique <- index_scores_elix %>% 
  select(patient_num, elixhauser_score) %>% 
  left_join(demo_df, by = 'patient_num')
length(unique(index_scores_elix$patient_num))
```

## Elixhauser comorbidity matrix

Heatmap of number of patients in each comorbidity type:

```{r}
n_comorbs <- colSums(index_scores_elix[, comorbs_elix])
pos_comorbs <- names(n_comorbs[n_comorbs > 0])
elix_mat <- cor(index_scores_elix[, pos_comorbs])
heatmap(elix_mat, cexRow = 0.5, cexCol = 0.5)
elix_obfus_table <- elix_stats(comorb_unique, blur_abs, mask_thres)

other_obfus_table <-
  bind_rows(
    nstay_obfus_table,
    readmission_obfus_table,
    severity_obfus_table,
    survival_obfus_table,
    elix_obfus_table
  )
datatable(other_obfus_table, options = list(pageLength = 14))
```

### Elixhauser Distribution

```{r}
comorb_unique %>% 
  mutate(severe = as.factor(severe) %>% 
           fct_recode('Severe' = '1', 'Non-severe' = '0')) %>% 
  ggplot(aes(x = neuro_post, y = elixhauser_score, fill = neuro_post)) + 
  geom_boxplot(outlier.colour = NA) + 
  geom_jitter(alpha = 0.2, width = 0.2, size = 0.6) +
  guides(fill = FALSE) +
  labs(x = NULL) +
  NULL
```

### Elixhauser Comorbidity Table

```{r}
scores_unique <- left_join(index_scores_elix, demo_df, by = 'patient_num')

scores_neuro <- obs_raw %>% 
  # 1 patient can have different code but each only counted once
  distinct(patient_num, concept_code) %>% 
  left_join(neuro_icds_10, by = c('concept_code' = 'icd')) %>% 
  left_join(scores_unique, by = 'patient_num') %>% 
  filter(!is.na(elixhauser_score)) %>%
  mutate(concept_code = case_when(
           is.na(`Neurological Disease Category`) ~ 'NN',
           TRUE ~ concept_code) %>% 
           as.factor() %>% 
           fct_reorder(- elixhauser_score),
         `Neurological Disease Category` =
           as.factor(`Neurological Disease Category`) %>% 
           fct_reorder(elixhauser_score)) %>% 
  {.}

elix_obfus_table1 <- 
  Reduce(function(...) left_join(..., by = c("Comorbidity", "Abbreviation")), 
       lapply(c('no_neuro_cond', 'neuro_cond'), list_table1, 
              df = scores_unique,
              num_pats = nrow(demo_df),
              comorb_names = comorb_names_elix,
              blur_abs = blur_abs, mask_thres = mask_thres)) %>% 
  mutate(
    n_Total = n_no_neuro_cond + n_neuro_cond,
    prop_Total = n_Total / nrow(demo_raw)
  ) %>%
  arrange(desc(n_Total))

elix_obfus_table1 %>% 
  transmute(
    Comorbidity, Abbreviation,
    `Has neurological condition` = concat(n_neuro_cond, prop_neuro_cond),
    `No neurological condition` = concat(n_no_neuro_cond, prop_no_neuro_cond),
    Total = concat(n_Total, prop_Total)) %>% 
  datatable() 

elix_obfus_table1 %>% 
  select(- contains('prop_'), - n_Total) %>% 
  pivot_longer(contains('n_')) %>% 
  mutate(fullname = glue::glue('{Comorbidity} ({Abbreviation})')) %>% 
  ggplot(aes(x = value, y = fct_reorder(fullname, value), fill = name)) +
  geom_col(position = 'stack') +
  scale_fill_brewer(palette = 'Dark2', direction = -1,
                    guide = guide_legend(reverse=TRUE)) +
  labs(y = NULL)
```

## Short term outcomes

Do hospitalized COVID-19 patients with neurological conditions have a different short-term natural history (hospital days, death, severity, readmission) from those without?

Resolution: Elixhauser score vs comorbidity type, Neurological condition (TRUE/FALSE) vs. neurological ICD code?

### Severity, Death, Length of stay

```{r}
severe_reg_elix <- glm(severe ~ neuro_post + elixhauser_score + sex + age_group + race,
   data = scores_unique, family = 'binomial')
severe_reg_elix
deceased_reg_elix <- glm(deceased ~ neuro_post + elixhauser_score + sex + age_group + race,
   data = scores_unique, family = 'binomial')
deceased_reg_elix

n_stay_reg_elix <- lm(n_stay ~ neuro_post + elixhauser_score + race + sex + age_group, 
   data = scores_unique)

n_stay_reg_elix
```

**Elixhauser**

```{r}
nstay_df <- neuro_patients %>% 
  bind_rows(non_neuro_patients) %>% 
  left_join(demo_df, by = 'patient_num') %>%
  left_join(index_scores_elix, by = 'patient_num') %>%
  # mutate(concept_code = fct_reorder(concept_code, n_stay)) %>% 
  left_join(neuro_icds_10, by = c('concept_code' = 'icd')) %>%
  mutate(full_icd = case_when(
    concept_code == 'NN' ~ 'No neurological condition',
    TRUE ~ paste0(`ICD-10 Description`, ' (', concept_code, ')')) %>%
             as.factor() %>% fct_reorder(n_stay)) 

summarised_obfus_icd <- nstay_df %>%
  group_by(concept_code) %>% 
  summarise(mean_stay = mean(n_stay),
            median_stay = median(n_stay),
            sd_stay = sd(n_stay),
            mean_elix = mean(elixhauser_score, na.rm = TRUE),
            median_elix = median(elixhauser_score, na.rm = TRUE),
            sd_elix = sd(elixhauser_score, na.rm = TRUE),
            n_patients = n(),
            prop_deceased = mean(deceased),
            prop_severe = mean(severe),
            .groups = 'drop') %>% 
  blur_it('n_patients', blur_abs, mask_thres)

```

```{r save-results}
list_results = list(
  demo_table = demo_obfus_table,
  other_obfus_table = other_obfus_table,
  elix_obfus_table1 = elix_obfus_table1,
  summarised_obfus_icd = summarised_obfus_icd
)

list_results <- lapply(list_results, function(x) mutate(x, site = mysite))
list_results <- c(
  list_results,
  list(
    site = mysite,
    elix_mat = elix_mat,
    n_stay_reg_elix = n_stay_reg_elix,
    severe_reg_elix = severe_reg_elix,
    deceased_reg_elix = deceased_reg_elix
  )
)

site_results <- paste0(mysite, '_results')
assign(site_results, list_results)
save(list = site_results,
     file = paste0('results/', mysite, '-results.rda'))
```
