---
title: "Neurological condition analysis"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "htmls") })
---

```{r setup, message=FALSE, warning=FALSE, results='hide'}
my_packages <- c('tidyverse', 'readxl', 'DT', 'rcartocolor', 'gghighlight', 
                 'cowplot', 'broom', 'devtools', 'tableone')

existing_pkgs <- installed.packages()[,"Package"]
to_install <- setdiff(my_packages, existing_pkgs)
if (length(to_install)) install.packages(to_install)
if (!'icd' %in% existing_pkgs) devtools::install_github('jackwasey/icd')

lapply(my_packages, library, character.only = TRUE)
theme_set(theme_bw() + 
            theme(legend.title = element_blank(),
                  panel.grid.minor = element_blank()))
for (r_file in list.files('R/', full.names = TRUE)) source(r_file)
set.seed(1)
```

```{r penn, message=FALSE}
mysite = 'penn'
mask_thres = 10 # count mask threshold
blur_abs = 3 # absolute max of blurring range

clin_raw <-
 read_csv(
   '../thrombotic-penn/data/LocalPatientClinicalCourse.csv',
   col_types = list(patient_num = col_character())
 )
demo_raw <-
 read_csv(
   '../thrombotic-penn/data/LocalPatientSummary.csv',
   col_types = list(patient_num = col_character()),
   na = '1900-01-01'
 )
obs_raw <-
 read_csv(
   '../thrombotic-penn/data/LocalPatientObservations.csv',
   col_types = list(patient_num = col_character())
 )
```


```{r nwu, message=FALSE}
# clin_raw <- read_csv(
#   '../../Phase2.1/Date_Shifted_Files/LocalPatientClinicalCourse_Shift.csv',
#   col_types = list(patient_num = col_character()))
# demo_raw <- read_csv(
#   '../../Phase2.1/Date_Shifted_Files/LocalPatientSummary_Shift.csv',
#   col_types = list(patient_num = col_character()),
#   na = '1900-01-01')
# obs_raw <- read_csv('../../Phase2.1/LocalPatientObservations.csv',
#                     col_types = list(patient_num = col_character()))
```


```{r message=FALSE, warning=FALSE}
loinc <- read_csv('public-data/loinc-map.csv')
neuro_icds_10 <- read_excel('public-data/2020-09-10_neuro-icd10.xlsx') %>% 
  rename('icd' = `ICD-10`)
```

## Data pre-processing

In this initial analysis, we will consider neuro patients as those with a neuro code on the day **of** or **anytime after** their first day of admission.
These patients developed neurological conditions after COVID-19 related admission.

Only examining ICD-10 for now.

*Note: We will need to discuss whether we want to include codes assigned on the day of admission as comorbidities.*


```{r}
neuro_patients <- obs_raw %>% 
  filter(
    days_since_admission >= 0,
    concept_code %in% neuro_icds_10$icd) %>% 
  distinct(patient_num, concept_code)

neuro_pt_post <- unique(neuro_patients$patient_num)
  
non_neuro_patients <- data.frame(
  patient_num = setdiff(demo_raw$patient_num, neuro_pt_post)) %>% 
  mutate(concept_code = 'NN')

readmissions <- clin_raw %>% 
  group_by(patient_num) %>% 
  mutate(delta_hospitalized = diff(c(in_hospital[1], in_hospital))) %>% 
  ungroup() %>% 
  filter(delta_hospitalized != 0,
         in_hospital == 1) %>%
  add_count(patient_num, name = 'n_readmissions') %>%
  arrange(desc(n_readmissions))

length(unique(readmissions$patient_num))

re_admin_stats <- readmissions %>% 
  select(patient_num, n_readmissions) %>% 
  distinct() %>% 
  count(n_readmissions)

re_admin_stats

re_admin_stats %>% 
  ggplot() +
  geom_col(aes(x = n_readmissions, y = n))
```

Length of stay is estimated based on the last day of having an "observation".

```{r}
days_count_min_max <- obs_raw %>%
  group_by(patient_num) %>%
  summarise(
    distinct_days = n_distinct(days_since_admission),
    min_hos = min(days_since_admission),
    # n_stay = max(days_since_admission),
    .groups = 'drop'
  )

demo_df <- demo_raw %>%
  mutate(time_to_severe = severe_date - admission_date,
         time_to_severe = ifelse(time_to_severe < 0, NA, time_to_severe),
         time_to_death = death_date - admission_date,
         time_to_death = ifelse(time_to_death < 0, NA, time_to_death),
         readmitted = patient_num %in% readmissions$patient_num,
         neuro_post = patient_num %in% neuro_pt_post %>% 
           as.factor() %>% 
           fct_recode(`Has neurological condition` = "TRUE",
                      `No neurological condition` = "FALSE"),
         Survival = as.factor(deceased) %>% 
           fct_recode(Alive = "0", Deceased = "1"),
         sex = as.factor(sex),
         race = as.factor(race),
         age_group = as.factor(age_group),
         Severity = as.factor(severe) %>% 
           fct_recode(Severe = "1", `Non-severe` = "0"),
         n_stay = as.numeric(last_discharge_date - admission_date,
                             units="days")) %>% 
  left_join(days_count_min_max, by = 'patient_num')
```

## Demographics (Table 1)

```{r}
formatted_demo <- demo_df %>% 
  select(Sex = sex, `Age group` = age_group, 
         Race = race, Severity, Survival, 
         `Time to severity onset (days)` = time_to_severe, 
         `Time to death (days)` = time_to_death, 
         Readmitted = readmitted, neuro_post)

## export table 1 to csv
table_one_vars <- setdiff(colnames(formatted_demo), 'neuro_post')
demo_table <- tableone::CreateTableOne(
  data = formatted_demo,
  vars = table_one_vars,
  strata= "neuro_post")

table_one <- print(
  demo_table,
  showAllLevels = TRUE,
  formatOptions = list(big.mark = ","),
  printToggle = FALSE
)

## print table 1 html
table1::table1(
  ~ Sex + `Age group` + Race + Severity + Survival + Readmitted +
    `Time to severity onset (days)` + `Time to death (days)` 
  | neuro_post,
  data = formatted_demo,
  overall = "Total",
  export = "results/demographics.csv")
```

## Comorbidity

```{r}
n_patients <- nrow(demo_raw)
obs_raw <- obs_raw %>% 
  filter(concept_type %in% c("DIAG-ICD10", "DIAG-ICD9"))
table(demo_df$neuro_post)
```

New variable `neuro_post` to indicate whether the patient developed neurological conditions after COVID-19 related admission.

## Map comorbidities

List of comorbidity abbreviations for reference purposes,
then run the `map_char_elix_codes()` function:

```{r}
# for elixhauser
comorb_names_elix <- get_quan_elix_names()
comorbs_elix <- as.vector(comorb_names_elix$Abbreviation)

# t1: earliest time point to consider comorbidities
# t2: latest time point to consider comorbidities
# example <- t1 = -365, and t2 = -1 will map all all codes up to a year prior but before admission (admission = day 0)

comorb_elix <- map_char_elix_codes(
  df = obs_raw,
  comorb_names = comorb_names_elix,
  t1 = -365,
  t2 = -15,
  map_type = 'elixhauser'
)
```

`map_char_elix_codes()` outputs a 2-element lists: 

<!-- 1- Dataframe indicating which comorbidities were mapped -->

1- Dataframe of the Charlson/Elixhauser comorbidity index scores

2- Dataframe of the specific mapped codes arranged by frequency.
Comorbidity mapping of all patients with ICD codes prior to admission, regardless of whether or not they had a comorbidity.

```{r}
# elixhauser
index_scores_elix <- comorb_elix$index_scores %>% 
  rename('elixhauser_score' = van_walraven_score)
# van Walraven is a modification of Elixhauser comorbidity measure
# doi.org/10.1097/MLR.0b013e31819432e5
mapped_codes_table_elix <- comorb_elix$mapped_codes_table
comorb_names_elix$Abbreviation <- as.character(comorb_names_elix$Abbreviation)
```

Here we can determine how many unique patients we have comorbidity index for prior to admission:
```{r}
# elixhauser
length(unique(index_scores_elix$patient_num))
comorb_unique <- select(index_scores_elix, patient_num, elixhauser_score)
```

## Elixhauser comorbidity matrix

Heatmap of number of patients in each comorbidity type:

```{r}
n_comorbs <- colSums(index_scores_elix[, comorbs_elix])
pos_comorbs <- names(n_comorbs[n_comorbs > 0])

index_scores_elix[, pos_comorbs] %>% 
  cor() %>% 
  heatmap(cexRow = 0.5, cexCol = 0.5)
```

### Elixhauser Distribution

```{r}
comorb_unique %>% 
  left_join(demo_df, by = 'patient_num') %>% 
  mutate(severe = as.factor(severe) %>% 
           fct_recode('Severe' = '1', 'Non-severe' = '0')) %>% 
  ggplot(aes(x = neuro_post, y = elixhauser_score, fill = neuro_post)) + 
  geom_boxplot(outlier.colour = NA) + 
  geom_jitter(alpha = 0.2, width = 0.2, size = 0.6) +
  guides(fill = FALSE) +
  labs(x = NULL) +
  NULL
```

### Elixhauser Comorbidity Table

```{r}
concat <- function(x, y){
  paste0(x, ' (', round(y, 3)*100, '%)')
}

scores_unique <- index_scores_elix %>% 
  left_join(demo_df, by = 'patient_num') %>% 
  # mutate(severe = as.factor(severe) %>% 
  #          fct_recode('Severe' = '1', 'Non-severe' = '0')) %>% 
  {.}

scores_neuro <- obs_raw %>% 
  # 1 patient can have different code but each only counted once
  distinct(patient_num, concept_code) %>% 
  left_join(neuro_icds_10, by = c('concept_code' = 'icd')) %>% 
  left_join(scores_unique, by = 'patient_num') %>% 
  filter(!is.na(elixhauser_score)) %>%
  mutate(concept_code = case_when(
           is.na(`Neurological Disease Category`) ~ 'NN',
           TRUE ~ concept_code) %>% 
           as.factor() %>% 
           fct_reorder(- elixhauser_score),
         `Neurological Disease Category` =
           as.factor(`Neurological Disease Category`) %>% 
           fct_reorder(elixhauser_score)) %>% 
  {.}

combined_table1_elix <- get_table1(
  index_scores_elix %>% filter(patient_num %in% neuro_pt_post),
  comorbidities = comorb_names_elix,
  comorbidities_map = comorb_names_elix$Abbreviation) %>% 
  rename('n_neuro_pats' = n_patients) %>% 
  left_join(get_table1(index_scores_elix,
                       comorbidities_map = comorb_names_elix$Abbreviation, 
                       comorbidities = comorb_names_elix), 
            by = c("Comorbidity", "Abbreviation"))

## apply blurring and masking
blur_vec <- matrix(sample(seq(-blur_abs, blur_abs),
                   4*length(comorbs_elix),
                   replace = TRUE), ncol = 4)

combined_table1_elix <- combined_table1_elix %>%
  mutate(
    n_patients = ifelse(n_patients < mask_thres, 0, n_patients + blur_vec[, 1]),
    n_neuro_pats = ifelse(n_neuro_pats < mask_thres, 0, n_neuro_pats + blur_vec[, 2]),
    n_non_neuro = n_patients - n_neuro_pats,
    n_non_neuro = ifelse(abs(n_non_neuro) < mask_thres, 0, n_non_neuro + blur_vec[, 3]),
    prop_patients = n_patients / nrow(demo_raw),
    prop_neuro_pats = n_neuro_pats / nrow(demo_raw),
    prop_non_neuro = n_non_neuro / nrow(demo_raw)
  ) %>%
  arrange(desc(n_neuro_pats))

combined_table1_elix %>% 
  transmute(
    Comorbidity, Abbreviation,
    `Has neurological condition` = concat(n_neuro_pats, prop_neuro_pats),
    `No neurological condition` = concat(n_non_neuro, prop_non_neuro),
    Total = concat(n_patients, prop_patients)) %>% 
  datatable() 

combined_table1_elix %>% 
  pivot_longer(c(n_patients, n_neuro_pats)) %>% 
  mutate(fullname = glue::glue('{Comorbidity} ({Abbreviation})'))%>% 
  ggplot(aes(x = value, y = fct_reorder(fullname, value), fill = name)) +
  geom_col(position = 'dodge') +
  scale_fill_brewer(palette = 'Dark2', direction = -1,
                    guide = guide_legend(reverse=TRUE),
                    label = c('Neuro', 'All')) +
  labs(y = NULL)
```

## Short term outcomes

Do hospitalized COVID-19 patients with neurological conditions have a different short-term natural history (hospital days, death, severity, readmission) from those without?

Resolution: Elixhauser score vs comorbidity type, Neurological condition (TRUE/FALSE) vs. neurological ICD code?

### Severity, Death, Length of stay

```{r}
severe_reg_elix <- glm(severe ~ neuro_post + elixhauser_score + sex + age_group + race,
   data = scores_unique, family = 'binomial') %>% 
  tidy()
severe_reg_elix
deceased_reg_elix <- glm(deceased ~ neuro_post + elixhauser_score + sex + age_group + race,
   data = scores_unique, family = 'binomial') %>% 
  tidy()
deceased_reg_elix

n_stay_reg_elix <- lm(n_stay ~ neuro_post + elixhauser_score + race + sex + age_group, 
   data = scores_unique) %>%
  tidy()

n_stay_reg_elix
```


## Comorbidities by Diagnosis

Here we will use the `obs_processed` dataframe in contrast to our `scores_unique` dataframe as the former is in long format

**Elixhauser**

```{r}
scores_neuro %>% 
  distinct(patient_num, concept_code, `Neurological Disease Category`) %>%
  left_join(index_scores_elix %>% select(patient_num, elixhauser_score), by = "patient_num") %>%
  filter(!is.na(elixhauser_score)) %>% 
  ggplot(aes(y = concept_code, x = elixhauser_score, fill = concept_code)) + 
  facet_grid(rows = vars(`Neurological Disease Category`),
             scales = 'free', space = 'free') +
  geom_boxplot(alpha = 0.6) + 
  scale_fill_viridis_d(guide = FALSE, direction = -1) +
  labs(x = 'Elixhauser score', y = NULL) +
  NULL +
  theme(panel.grid = element_blank(),
        legend.title = element_text(),
        strip.text.y = element_text(angle = 0),
        panel.spacing.x = unit(10, "points"),
        panel.spacing.y = unit(0, "points"))
```


## Evaluating Timeframe of Comorbidities

Qs: When are these comorbidities getting mapped. Understanding ICD code assignnment can help us prevent confusing complications with comorbidities and vice versa. 

It would be helpful to review the literature as well, what did the original developers of the comorbidity indexes do? How were they validated?
```{r eval=FALSE, include=FALSE}
# maps codes that occurred anytime 356 days prior to COVID admission
multi_comorb <- function(x, ...) {
  process_tables(
    map_char_elix_codes(obs_raw, comorb_names_elix, x$t1, x$t2, 
                       map_type = 'elixhauser')[['index_scores']],
    ...) %>%
    mutate(time_map = x$label) 
}

time_df <- tribble(
  ~ t1, ~ t2, ~ label,
  -365, -1, "1 to 365 days prior admission",
  -365, -15, "15 to 365 days prior admission",
  -14, -1, "14 to 1 days prior admission",
  0, 0, "Only Day of Admission",
  1, 1000, "Any Day After Admission"
)

all_tables <- lapply(time_df %>% split(1:5), multi_comorb, 
                     comorbidities = comorb_names_elix) %>%
  do.call(rbind, .) %>% 
  mutate(time_map = fct_relevel(time_map, time_df$label))
```

### Timeline - All Patients
```{r eval=FALSE, include=FALSE}
ggplot(all_tables, aes(x = time_map,  y = n_patients, fill = time_map)) +
  geom_bar(stat = "identity") +
  facet_wrap( ~ Abbreviation) +
  scale_fill_carto_d() +
  theme(
    legend.position = c(0.8, 0.12),
    legend.key.height = unit(3, 'mm'),
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
```

```{r}
nstay_df <- neuro_patients %>% 
  bind_rows(non_neuro_patients) %>% 
  left_join(demo_df, by = 'patient_num') %>%
  # left_join(scores_unique, by = 'patient_num') %>%
  # mutate(concept_code = fct_reorder(concept_code, n_stay)) %>% 
  left_join(neuro_icds_10, by = c('concept_code' = 'icd')) %>%
  mutate(full_icd = case_when(
    concept_code == 'NN' ~ 'No neurological condition',
    TRUE ~ paste0(`ICD-10 Description`, ' (', concept_code, ')')) %>%
             as.factor() %>% fct_reorder(n_stay)) 
  

summarised_stay <- nstay_df %>%
  group_by(concept_code) %>% 
  summarise(mean_stay = mean(n_stay, na.rm = TRUE),
            median_stay = median(n_stay, na.rm = TRUE),
            sd_stay = sd(n_stay, na.rm = TRUE),
            n_patients = n(),
            n_patients = ifelse(n_patients < mask_thres, 0, n_patients + blur_vec),
            prop_deceased = mean(deceased, na.rm = TRUE),
            prop_severe = mean(severe, na.rm = TRUE),
            .groups = 'drop')

summarised_stay
```

```{r save-results}
list_results = list(
  summarised_stay = summarised_stay, 
  n_stay_reg_elix = n_stay_reg_elix, 
  severe_reg_elix = severe_reg_elix, 
  deceased_reg_elix = deceased_reg_elix, 
  # table_one = table_one,
  combined_table1_elix = combined_table1_elix)

site_results <- paste0(mysite, '_results')
assign(site_results, list_results)
save(list = site_results,
     file = paste0('results/', mysite, '-results.rda'))

# write.csv(table_one, file = 'results/penn-table1.csv')
# write.csv(combined_table1_elix,
#           paste0('results/', mysite, '-combined-elix.csv'), 
#           row.names = FALSE)
```

```{r nwu-save-results}
# summarised_stay_nwu = summarised_stay
# n_stay_reg_nwu = n_stay_reg
# severe_reg_nwu = severe_reg
# deceased_reg_nwu = deceased_reg
# table_one_nwu = table_one
# write.csv(table_one_nwu, file = "results/nwu-table1.csv")
# 
# save(summarised_stay_nwu,
#      severe_reg_nwu,
#      deceased_reg_nwu,
#      table_one_nwu,
#      n_stay_reg_nwu, 
#      file = 'results/nwu-results.rda')
```

