---
title: "Neuroanalysis"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(gghighlight)

theme_set(theme_bw() + 
            theme(legend.title = element_blank(),
                  panel.grid.minor = element_blank()))
for (r_file in list.files('R/', full.names = TRUE)) source(r_file)
```

```{r message=FALSE}
# labs_raw <- read_csv('../thrombotic-penn/data/labs_long_thrombo_v2.csv')
clin_raw <- read_csv('../thrombotic-penn/data/LocalPatientClinicalCourseSHIFTED-UPenn.csv')
demo_raw <- read_csv('../thrombotic-penn/data/LocalPatientSummarySHIFTED-UPenn.csv')
obs_raw <- read_csv('../thrombotic-penn/data/LocalPatientObservations-UPenn.csv')
loinc <- read_csv('public-data/penn-loinc-map.csv')
neuro_icds_10 <- read_excel('public-data/2020-09-10_neuro-icd10.xlsx') %>% 
  rename('icd' = `ICD-10`)

neuro_patients <- obs_raw %>% 
  filter(concept_code %in% neuro_icds_10$icd) %>% 
  distinct(patient_num, concept_code)

non_neuro_patients <- data.frame(
  patient_num = setdiff(demo_raw$patient_num, unique(neuro_patients$patient_num))) %>% 
  mutate(concept_code = 'NN')

readmissions <- clin_raw %>% 
  group_by(patient_num) %>% 
  mutate(delta_hospitalized = diff(c(in_hospital[1], in_hospital))) %>% 
  ungroup() %>% 
  filter(delta_hospitalized != 0) %>% 
  add_count(patient_num, name = 'status_changes') %>% 
  filter(status_changes > 1) %>% 
  mutate(n_readmissions = (status_changes - 1)/2) %>% 
  arrange(desc(n_readmissions))

length(unique(readmissions$patient_num))

re_admin_stats <- readmissions %>% 
  select(patient_num, n_readmissions) %>% 
  distinct() %>% 
  count(n_readmissions)

re_admin_stats %>% 
  ggplot() +
  geom_col(aes(x = n_readmissions, y = n))

```

Length of stay is estimated based on the last day of having an "observation".

```{r}
days_count_min_max <- obs_raw %>%
  group_by(patient_num) %>%
  summarise(
    distinct_days = n_distinct(days_since_admission),
    min_hos = min(days_since_admission),
    n_stay = max(days_since_admission),
    .groups = 'drop'
  )

demo_df <- demo_raw %>% 
  mutate(time_to_severe = severe_date_shift - admission_date_shift,
         time_to_severe = ifelse(time_to_severe < 0, NA, time_to_severe),
         time_to_death = death_date_shift - admission_date_shift,
         time_to_death = ifelse(time_to_death < 0, NA, time_to_death),
         readmitted = patient_num %in% readmissions$patient_num,
         neuro_cond = as.factor(patient_num %in% neuro_patients$patient_num) %>% 
           fct_recode(`Has neurological condition` = "TRUE",
                      `No neurological condition` = "FALSE"),
         Survival = as.factor(deceased) %>% 
           fct_recode(Alive = "0", Deceased = "1"),
         sex = as.factor(sex),
         race = as.factor(race),
         age_group = as.factor(age_group),
         Severity = as.factor(severe) %>% 
           fct_recode(Severe = "1", `Non-severe` = "0"),
         ) %>% 
  left_join(days_count_min_max, by = 'patient_num')

nat_hist_df <- clin_raw %>% 
  group_by(patient_num) %>% 
  summarise(ever_severe = max(severe) %>% as.logical(), 
            death = max(deceased) %>% as.logical(),
            .groups = 'drop') %>% 
  left_join(demo_df %>% select(patient_num, readmitted), 
            by = 'patient_num')

```

## Demographics (Table 1)

```{r}
demo_table <- table1::table1(
  ~ Sex + `Age Group` + Race + Severity + 
    Survival + `Time to severity onset (days)` + `Time to death (days)` 
  | neuro_cond,
  data = demo_df %>% 
  select(patient_num, Sex = sex, `Age Group` = age_group, 
         Race = race, Severity, Survival, 
         `Time to severity onset (days)` = time_to_severe, 
         `Time to death (days)` = time_to_death, readmitted, neuro_cond),
  overall = "Total",
  export = "results/demographics.csv")
demo_table
```

```{r eval=FALSE, include=FALSE}
# penn_labs <- obs_raw %>% 
#   left_join(loinc, by = c('concept_code' = 'LOINC')) %>% 
#   pivot_wider(id_cols = c(patient_num, days_since_admission), 
#               names_from = short_name, values_from = value,
#               values_fn = median) %>%
#   arrange(patient_num, days_since_admission) %>% 
#   rownames_to_column('id') %>%
#   rename('covid_id' = patient_num, 'time' = days_since_admission) %>% 
#   mutate(id = as.integer(id)) %>% 
#   {.}
# non_lab_value_names <- c('id', 'covid_id', 'time')
# lab_value_names <- setdiff(names(penn_labs), non_lab_value_names)
```

## Short term

Do hospitalized COVID-19 patients with neurological conditions have a different short-term natural history (hospital days, death, severity, readmission) from those without?

### Length of stay

```{r}
demo_df %>% 
  ggplot(aes(x = neuro_cond, y = n_stay, fill = neuro_cond, color = neuro_cond)) +
  geom_dotplot(binaxis = "y", stackdir = "centerwhole", binwidth = 1) +
  scale_fill_discrete(guide = FALSE) +
  scale_color_discrete(guide = FALSE)

demo_df %>% 
  ggplot(aes(x = neuro_cond, y = n_stay, fill = neuro_cond)) +
  geom_boxplot() +
  scale_fill_discrete(guide = FALSE)

lm(n_stay ~ neuro_cond + race
   # + sex + age_group
   , data = demo_df) %>% 
  summary()
```

NN: non-neurological code

```{r}
neuro_patients %>% 
  bind_rows(non_neuro_patients) %>% 
  left_join(demo_df, by = 'patient_num') %>% 
  mutate(concept_code = fct_reorder(concept_code, n_stay)) %>% 
  ggplot(aes(x = concept_code, y = n_stay, fill = concept_code)) +
  labs(x = NULL) +
  geom_boxplot() + 
  scale_fill_viridis_d(guide = FALSE) +
  gghighlight(concept_code != 'NN') +
  NULL
```

### Death, severity, readmission

```{r warning=FALSE, fig.width=10}
run_fish('death')
run_fish('ever_severe')
run_fish('readmitted')
```

TODO : add race, sex, age_group
```{r}
demo_df %>% 
  left_join(nat_hist_df, by = 'patient_num') %>% 
  select(neuro_cond, death) %>% 
  table()

demo_df %>% 
  left_join(nat_hist_df, by = 'patient_num') %>% 
  select(neuro_cond, ever_severe) %>% 
  table()

demo_df %>% 
  select(neuro_cond, readmitted) %>% 
  table()

summarise_nat_hist <- function(variable){
  demo_df %>% 
  left_join(nat_hist_df, by = c('patient_num', 'readmitted')) %>% 
  lm(as.formula(paste(variable, '~ neuro_cond')), data = .) %>% 
    broom::tidy() %>% 
    mutate(dependent_var = variable)
}

bind_rows(summarise_nat_hist('death'),
          summarise_nat_hist('ever_severe'),
          summarise_nat_hist('readmitted'))
```

